Exploring a pragmatic account of Zipf's Law of Abbreviation using Bayesian data analysis

### 2019-01-08 paper/paper.Rmd
additional tweaks before ending work today

---
title: "A pragmatic account of Zipf's Law of Abbreviation via Bayesian data analysis"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

## Rational speech act theory as a model for pragmatic participant behavior

@@

\begin{equation}
S_i(u|m) \propto e^{U(u;m)}
\end{equation}
Where
\begin{equation}
U(u;m) = \alpha(log(L_{i-1}(m|u)) - cost(u))
\end{equation}
And 
\begin{equation}
L_{i-1}(m|u) \propto S_{i-1}(u|m)\times p(m)
\end{equation}
Defining nested speaker and listener agents could, in principle lead to infinite recursion. We take a *literal listener*, denoted $L_0(m|u) as a basecase. The literal listener does not reason about a speaker model, rather this agent considers the literal semantics of the utterance.
\begin{equation}
L_0(m|u) \propto \delta_u(m)\times p(m)
\end{equation}
Where
\begin{equation}
\delta_u(m)=
\begin{cases}
1, & \text{if } m \in [[u]]\\
0, & o.w.
\end{cases}
\end{equation}

# Modeling framework

@@

```{r plate-notation, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
img <- png::readPNG("figs/plate-temp.png")
grid::grid.raster(img)
```

For each subject we use $n=200$ samples using Metropolis-Hasting to approximate the posterior distribution over parameters.

Figure X displays inferred subject params

**NEED SOME FIGURES ON THE INFERRED PARAMS**

@@

```{r plot-inferred-params, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/inferred-params-plot.png")
grid::grid.raster(img)
```

@@

```{r plot-post-predictives, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/bda-vs-baseline-plot.png")
grid::grid.raster(img)
```

>>>>

---
title: "Exploring a pragmatic account of Zipf's Law of Abbreviation using Bayesian data analysis"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

# Rational speech act theory as a model for pragmatic participant behavior

@@

\begin{equation}
S_i(u|m) \propto e^{-\alpha\times U(u;m)}
\end{equation}
Where
\begin{equation}
U(u;m) = -log(L_{i-1}(m|u)) - cost(u)
\end{equation}
And 
\begin{equation}
L_{i-1}(m|u) \propto S_{i-1}(u|m)\times p(m)
\end{equation}
Defining nested speaker and listener agents could, in principle lead to infinite recursion. We take a *literal listener*, denoted $L_0(m|u)$ as a base-case. The literal listener does not reason about a speaker model, rather this agent considers the literal semantics of the utterance.
\begin{equation}
L_0(m|u) \propto \delta_u(m)\times p(m)
\end{equation}
Where
\begin{equation}
\delta_u(m)=
\begin{cases}
1, & \text{if } m \in [[u]]\\
0, & o.w.
\end{cases}
\end{equation}

```{r plot-inferred-params, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "TEMP - NEEDS UPDATING Inferred subject-level parameter distributions. Moving clock-wise from the upper-right facet -- inferred alpha values for subjects, the recursive depth for each subject, the probability that subjects are paying attention during the experiment, the inferred cost / referrent frequency structure used by the subject."}
img <- png::readPNG("figs/inferred-params-plot.png")
grid::grid.raster(img)
```

## Bayesian data analysis with RSA linking function

@@

```{r plate-notation, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "THIS IS A PLACE-HOLD AND IS INCORRECT"}
img <- png::readPNG("figs/plate-temp.png")
grid::grid.raster(img)
```

For each subject we use $n=200$ samples using Metropolis-Hasting to approximate the posterior distribution over parameters. Figure X displays inferred subject params.

@@

@@

```{r plot-post-predictives, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Posterior predictive values lead to far better fit to human data than a baseline model. Horizontal axes display counts of short-form usage from subjects in Kanwal et al. (2017). Verticle axes display model predictions. The left facet displays predicted short-form usage under a baseline model that implements the 'optimal' solution for each condition. The right facet displays predicted short-form usage under our BDA model."}
img <- png::readPNG("figs/bda-vs-baseline-plot.png")
grid::grid.raster(img)
```
